Report

### 2020-09-16 report.Rmd
Add a proper statistical model

library(drake)
library(tidyverse)
library(magrittr)
library(cowplot)
#theme_set(theme_cowplot())
```

@@

- Questions:
  1. Are assembly and disassembly of networks symmetric processes ?

@@

```{r p-tmp-dyn, cache = FALSE}
loadd(temporal_dynamics_plot)

@@

@@

```{r p-bivar, cache = FALSE, fig.cap = bivar_cap, fig.height = 10}
loadd(net_dyn_lm)
loadd(net_dyn_lm_plot)

@@

```{r}
loadd(temporal_dynamics_coef)
loadd(data)
first_year_biomass <-
  data %>%
  filter(nb_year == 0) %>%
  select(station, biomass, rel_bm, log_bm, rel_log_bm) %>%
  pivot_longer(-station, names_to = "coeff", values_to = "first_year") %>%
  group_by(coeff) %>%
  mutate(com_size = ifelse(first_year < median(first_year, na.rm = TRUE), "little", "big"))
median_biomass <- 
  data %>%
  select(station, biomass, log_bm) %>%
  pivot_longer(-station, names_to = "coeff", values_to = "biomass") %>%
  group_by(station, coeff) %>%
  summarise(bm_med = median(biomass, na.rm = TRUE)) %>%
  ungroup()

@@

```{r}
data %>%
  group_by(station) %>%
  summarise(med_bm = median(biomass)) %>%
  ggplot(aes(x = med_bm)) +
  geom_histogram() +
  labs(x = "Median total biomass over time by station (g)")
```

@@

>>>>

library(drake)
library(tidyverse)
library(magrittr)
library(cowplot)
source("R/misc.R") # Define your custom code as a bunch of functions.
source_dir(get_mypath("R")) # Define your custom code as a bunch of functions.

#theme_set(theme_cowplot())
```

@@

- Questions:
  1. Is a decrease or a increase in biomass related to a network
     structure modification?
     1.1 Is a contant decrease/increase in biomass correspond the same dynamic
     in network? Or there are non linear dynamics ?
  2. Are assembly and disassembly of networks symmetric processes?

@@

@@

```{r}
loadd(net_dyn_lm)
loadd(net_dyn_lm_plot)
```

@@

```{r p-bivar, cache = TRUE, fig.cap = bivar_cap, fig.height = 10}

@@

```{r}
loadd(temporal_dynamics_coef)
loadd(temporal_dynamics_plot)
loadd(full_data2)
```

```{r}
first_year_biomass <-
  full_data2 %>%
  filter(nb_year == 0) %>%
  select(station, biomass, rel_bm, log_bm, rel_log_bm) %>%
  pivot_longer(-station, names_to = "coeff", values_to = "first_year") %>%
  group_by(coeff) %>%
  mutate(com_size = ifelse(first_year < median(first_year, na.rm = TRUE), "little", "big")) %>%
  ungroup()

median_biomass <- 
  full_data2 %>%
  select(station, biomass, log_bm) %>%
  pivot_longer(-station, names_to = "coeff", values_to = "biomass") %>%
  group_by(station, coeff) %>%
  summarise(bm_med = median(biomass, na.rm = TRUE)) %>%
  ungroup()
three_first_year_biomass <-
  full_data2 %>%
  filter(nb_year <= 2) %>%
  select(station, nb_year, biomass, rel_bm, log_bm, rel_log_bm) %>%
  pivot_longer(cols = biomass:rel_log_bm, names_to = "coeff", values_to = "first_year") %>%
  group_by(station, coeff) %>%
  summarise(first_3_year = median(first_year, na.rm = TRUE)) %>%
  group_by(coeff) %>%
  mutate(com_size = ifelse(first_3_year < median(first_3_year, na.rm = TRUE), "little", "big")) %>%
  ungroup()

@@

```{r}
full_data2 %>%
  group_by(station) %>%
  summarise(med_bm = median(biomass)) %>%
  ggplot(aes(x = med_bm)) +
  geom_histogram() +
  labs(x = "Median total biomass over time by station (g)")
```

##

@@

## Biomass classification

```{r}
loadd(rigal_classification)
myload(biomass_ts_sax, envir = globalenv(),dir = get_mypath("data"))

biomass_classif_rigal <- rigal_classification %>%
  filter(variable %in% c("biomass", "log_bm")) %>%
  unnest(classif) %>%
  mutate(station = as.character(station)) %>%
  select(station, variable,shape_class)

p_biomass_classif_rigal <- biomass_classif_rigal %>%
  ggplot(aes(x = shape_class)) +
  geom_bar() +
  facet_wrap(~variable) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r p-tmp-dyn, cache = TRUE}

## Temporal dynamics decrease/increase constant

### Log biomass

```{r, fig.height = 45}
st_decrease_increase_rigal <- biomass_classif_rigal %>%
  filter(shape_class %in% c("increase_constant", "decrease_constant"))
data_decrease_increase <- full_data2 %>%
  filter(station %in% st_decrease_increase_rigal$station)

log_bm_min_max <- map_dbl(c(min, max),
  ~.x(data_decrease_increase$log_bm, na.rm = TRUE))
nb_year_min_max <- map_dbl(c(min, max),
  ~.x(data_decrease_increase$nb_year, na.rm = TRUE))
temporal_dynamics_plot$plot %<>%
  map(., ~ .x + lims(y = log_bm_min_max, x = nb_year_min_max))

temporal_dynamics_plot %>%
  filter(variable == "log_bm", station %in% st_decrease_increase_rigal$station) %>%
  plot_grid(plotlist = .$plot, ncol = 3, labels = st_decrease_increase_rigal$station)
```

### Raw Biomass

```{r, fig.height = 45}
st_decrease_increase_rigal_biomass <- rigal_classification %>%
  filter(variable == "bm_std") %>%
  unnest(classif) %>%
  filter(shape_class %in% c("increase_constant", "decrease_constant"))

bm_min_max <- map_dbl(
  c(min, max),
  ~.x(data_decrease_increase$biomass, na.rm = TRUE))
nb_year_min_max <- map_dbl(
  c(min, max),
  ~.x(data_decrease_increase$nb_year, na.rm = TRUE))
temporal_dynamics_plot[temporal_dynamics_plot$variable == "biomass",]$plot %<>%
  map(., ~ .x + lims(y = bm_min_max, x = nb_year_min_max))

temporal_dynamics_plot %>%
  filter(variable == "biomass", station %in% st_decrease_increase_rigal_biomass$station) %>%
  plot_grid(plotlist = .$plot, ncol = 3, labels = st_decrease_increase_rigal_biomass$station)
```

## Station grouping

### Rigal station increase / decrease

Les stations ont moins de 40 kilos au départ ont l'air comparable en terme de
dynamique.

```{r}

station_summary_biomass <- full_data2 %>%
  group_by(station) %>%
  summarise(bm_med = median(biomass, na.rm = TRUE)) %>%
  left_join(
    filter(three_first_year_biomass, coeff == "biomass") %>%
      select(station, first_3_year)
    , by =
    "station") %>%
  left_join(
    filter(first_year_biomass, coeff == "biomass") %>%
      select(station, first_year)
      , by =
    "station")

p_rigal_decrease_increase_slope_st <- 
  st_decrease_increase_rigal_biomass %>%
  left_join(station_summary_biomass, by = "station") %>%
  select(station, shape_class, linear_slope, linear_intercept, first_3_year, bm_med, first_year) %>%
  pivot_longer(cols = first_3_year:first_year, names_to = "variable",
    values_to = "value") %>%
  ggplot(aes(y = linear_slope, x = value, color = shape_class)) +
  geom_point() +
  facet_wrap(~variable)
p_rigal_decrease_increase_slope_st
```

### Tendancies in rigal increase / decrease

```{r}
rigal_biomass_coeff_dynamics <- st_decrease_increase_rigal_biomass %>%
  select(station, linear_slope, strd_error) %>%
  left_join(temporal_dynamics_coef, by = "station") %>%
  filter(!coeff %in% get_biomass_var(), term == "slope") %>%
  select(station, coeff, estimate, std.error, linear_slope, bm_std_error = strd_error) %>%
  mutate(reg_weight = (abs(estimate / std.error) + abs(linear_slope / bm_std_error))/2)
rigal_biomass_coeff_dynamics %>%
  ggplot(aes(x = reg_weight)) +
  geom_histogram() +
  facet_wrap(~coeff)

rigal_biomass_coeff_dynamics %>%
  ggplot(aes(x = linear_slope, y = estimate, size = reg_weight, weight = reg_weight)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~coeff, scales = "free_y")
rigal_biomass_coeff_dynamics %>%
  ggplot(aes(x = linear_slope, y = estimate, size = reg_weight)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~coeff, scales = "free_y")
```

### With group

#### Median

```{r}
group_results <- rigal_biomass_coeff_dynamics %>%
  left_join(station_summary_biomass, by = "station") %>%
  pivot_longer(cols = bm_med:first_year, names_to = "bm_summary_type", values_to = "bm_summary") %>%
  group_by(bm_summary_type) %>%
  mutate(
    median_group = ifelse(bm_summary < median(bm_summary, na.rm = TRUE), "little", "big"),
    first_quartile = map_chr(bm_summary, function (x, y) {
      if (is.na(x)) return(NA)

      if (x < y[1])
	output <- "first_quartile"
      else if (x > y[2])
	output <- "third_quartile"
      else
	output <- NA
      return(output)
}, y = quantile(bm_summary, probs = c(.25, .75), na.rm = TRUE))
  ) %>%
  ungroup()

##Summary of group 
group_results %>%
  ggplot(aes(x = bm_summary)) +
  geom_histogram() +
  facet_wrap(~bm_summary_type, scales = "free_x")

p1 <- group_results  %>%
  filter(bm_summary_type == "first_3_year") %>%
  ggplot(aes(x = linear_slope, y = estimate, color = median_group, size = reg_weight, weight = reg_weight)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~coeff, scales = "free_y")

p2 <- group_results  %>%
  filter(bm_summary_type == "first_3_year") %>%
  ggplot(aes(x = linear_slope, y = estimate, color = first_quartile, size = reg_weight, weight = reg_weight)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~coeff, scales = "free_y")

p3 <- group_results  %>%
  filter(bm_summary_type == "bm_med") %>%
  ggplot(aes(x = linear_slope, y = estimate, color = median_group, size = reg_weight, weight = reg_weight)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~coeff, scales = "free_y")

p1
p2
p3
```

#### Little and big rivers

```{r}
#grandes/petites rivières
myload(habitat_press, habitat_pressure, op_analysis, envir = environment(),
  dir = get_mypath("data"))
width_classif_mean <- habitat_press %>%
    mutate(width_classif_mean  = if_else(width_river_mean < mean(width_river_mean, na.rm = TRUE), "little",
      "big")) %>%
    select(station, width_river_mean, width_classif_mean)
habitat_press %>%
  ggplot(aes(x = width_river_mean)) +
  geom_histogram()

habitat_pressure %>%
  pivot_longer(cols = c(RC1, RC2), names_to = "PCA_axis", values_to = "value") %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(~PCA_axis)
protocol_station <- op_analysis %>%
  group_by(station) %>%
  summarise(protocol_type = unique(protocol_type))

habitat_press %>%
  left_join(protocol_station, by = "station") %>%
  ggplot(aes(x = width_river_mean, fill = protocol_type)) +
  geom_histogram()

group_results_stream_size <-
  group_results %>%
  left_join(
    habitat_pressure %>%
    mutate(RC1_med = if_else(RC1 < median(RC1, na.rm = TRUE), "little",
      "big")) %>%
    select(station, RC1, RC1_med),
  by = "station"
  ) %>%
  left_join(protocol_station, by = "station") %>%
  left_join(width_classif_mean, by = "station")

group_results_stream_size %>%
  filter(bm_summary_type == "first_3_year", linear_slope < 8) %>%
  ggplot(aes(x = linear_slope, y = estimate, color = protocol_type, size = reg_weight, weight = reg_weight)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~coeff, scales = "free_y") +
  labs(title = "Group as protocol type")

group_results_stream_size %>%
  filter(bm_summary_type == "first_3_year", linear_slope < 8) %>%
  ggplot(aes(x = linear_slope, y = estimate, color = protocol_type, size = reg_weight, weight = reg_weight)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~coeff, scales = "free_y") +
  labs(title = "Group as mean river size")

group_results_stream_size %>%
  filter(bm_summary_type == "first_3_year", linear_slope < 8) %>%
  ggplot(aes(x = linear_slope, y = estimate, color = width_classif_mean, size = reg_weight, weight = reg_weight)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~coeff, scales = "free_y") +
  labs(title = "Group as mean river size")

group_results_stream_size %>%
  filter(bm_summary_type == "bm_med", linear_slope < 8) %>%
  ggplot(aes(x = linear_slope, y = estimate, color = median_group, size = reg_weight, weight = reg_weight)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~coeff, scales = "free_y") +
  labs(title = "Group as mean biomass")

# Quelle est la compo des dynamique dans RC1
group_results_stream_size %>%
  filter(bm_summary_type == "first_3_year") %>%
  ggplot(aes(x = RC1)) +
  geom_histogram() +
  facet_wrap(~protocol_type)

group_results_stream_size %>%
  filter(bm_summary_type == "first_3_year") %>%
  ggplot(aes(x = bm_summary)) +
  geom_histogram() +
  facet_wrap(~protocol_type)
```

```{r}
model_int <- group_results_stream_size %>%
  filter(bm_summary_type == "first_3_year", linear_slope < 5000) %>%
  group_by(coeff) %>%
  nest() %>%
  mutate(model = map(data, ~lm(estimate ~ linear_slope * width_river_mean,.x)))

test <- model_int$model[[3]]
car::vif(test)

```

```{r}
group_results_stream_size %>%
  filter(bm_summary_type == "bm_med", linear_slope < 8) %>%
  ggplot(aes(x = linear_slope, y = estimate, color = median_group, size = reg_weight, weight = reg_weight)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~coeff, scales = "free") +
  labs(title = "Group as mean biomass")
```

```{r}
model_int <- group_results_stream_size %>%
  filter(bm_summary_type == "bm_med", linear_slope < 8) %>%
  mutate(
    increasing = linear_slope > 0,
    inc_f = as.factor(increasing) 
    ) %>%
  group_by(coeff) %>%
  nest() %>%
  mutate(
    mod1 = map(data, ~lm(estimate ~ linear_slope:(linear_slope <0) + linear_slope:(linear_slope >0),.x)),
    mod2 = map(data, ~lm(estimate ~ linear_slope:increasing,.x)),
    mod3 = map(data, ~lm(estimate ~ linear_slope:increasing:median_group,.x)),
    mod4 = map(data, ~lm(estimate ~ linear_slope:increasing:median_group, .x, weights = reg_weight)),
    mod5 = map(data, ~lm(estimate ~ linear_slope:inc_f:median_group, .x, weights = reg_weight)),
  )

model_out <- map(model_int$mod5, summary)
model_vif <- map(model_int$mod4, car::vif)
names(model_out) <- model_int$coeff 
model_out
```

```{r}
library(ggeffects)
mydf <- ggpredict(model_int$mod5[[1]], terms = c("linear_slope", "inc_f",
    "median_group"))
plot(mydf, add.data = TRUE)

plot_final_model <- function (ggpred = NULL, rawdata = NULL) {
  ggpred %<>%
    as_tibble %>%
    rename(increasing = group, bm_group = facet)

  ggpred %<>%
    filter(increasing == FALSE & x < 0 | increasing == TRUE & x > 0)

  if (is.null(rawdata)) {
  rawdata <- attr(ggpred, "rawdata")
  }

  p <- ggplot(ggpred, aes(x = x , y = predicted, color = bm_group)) +
    geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = bm_group), alpha =
      .1) +
    geom_line()
  p + geom_point(data = rawdata, aes(y = estimate, x = linear_slope, color =
      median_group))

}
```

```{r}
model_int %<>%
  mutate(
    pred5 = map(mod5, ~ggpredict(.x,
	terms = c("linear_slope [-8, -6, -4,-3, -2, -1, -.5, -.01, .01, .5, 1, 2,3, 4, 6]", "inc_f", "median_group"))),
    plot = map(pred5, ~plot(.x, add.data = TRUE)),
    myplot = map2(pred5, data, ~plot_final_model(ggpred = .x, rawdata = .y))
  )
plot_grid(plotlist = model_int$myplot, labels = model_int$coeff)
```

```{r}
library(segmented)

mod_tlvl <- lm(estimate ~ linear_slope, model_int$data[[2]])
mod_tlvl <- lm(
  estimate ~ linear_slope*decreasing*median_group,#(linear_slope < 0)*linear_slope + (linear_slope >= 0)*linear_slope,
  mutate(model_int$data[[2]], decreasing = linear_slope < 0)
)
seg_tlvl <- segmented(mod_tlvl, seg.Z=~linear_slope)
plot(seg_tlvl)

mod_richness <- lm(estimate ~ linear_slope, model_int$data[[2]])
test <- segmented(mod_richness, seg.Z=~linear_slope)
plot(test)

car::vif(test)
```

```{r}
#keep only complete protocol
group_results_stream_size %>%
  filter(protocol_type == "complete") %>%
  filter(bm_summary_type == "bm_med", linear_slope < 8) %>%
  ggplot(aes(x = linear_slope, y = estimate, color = median_group)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~coeff, scales = "free") +
  labs(title = "Group as mean biomass")

```

## Richness

```{r}
```

```{r}
st_big_bm_tendancies <- group_results %>%
  filter(abs(linear_slope) > 5000) %>%
  .$station %>%
  unique
plot_big_bm_tendancies <- temporal_dynamics_plot %>%
  filter(variable == "biomass", station %in% st_big_bm_tendancies) %>%
  .$plot
plot_grid(
  plotlist = map(plot_big_bm_tendancies, 
    ~.x + lims(x = c(0, 10), y = c(0, 155000)))
)
```

```{r}
rigal_shape_var <- rigal_classification %>%
  unnest(classif) %>%
  select(station, variable, shape_class) %>%
  mutate(station = as.character(station))

rigal_shape_var %>%
  filter(station %in% st_decrease_increase_rigal_biomass$station) %>%
  ggplot(aes(x = shape_class)) +
  geom_bar() +
  facet_wrap(~ variable) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}

```

## SAX

```{r, eval = FALSE}
library(jmotif)
loadd(full_data2)
sax_all_variable <- full_data2 %>%
  select(station, nb_year, biomass:log_bm) %>%
  pivot_longer(cols = biomass:log_bm, names_to = "variable", values_to = "value") %>%
  group_by(station, variable) %>%
  arrange(nb_year) %>%
  nest() %>%
  mutate(
    zts = map(data, function(x) znorm(x$value, threshold = 0.01)),
    paa = map(zts, ~paa(., 3)),
    sax = map_chr(paa, ~series_to_string(., 3))
  ) %>%
  ungroup()

# CBA:
sax_all_variable %>%
  filter(station %in% cba_rigal_sax$station, !variable %in% c("biomass", "nbnode")) %>%
  ggplot(aes(x = sax)) +
  geom_bar() +
  facet_wrap(facets = ~variable)

sax_all_variable %>%
  filter(
    station %in% cba_rigal_sax$station,
    !variable %in% c("biomass", "nbnode")
    ) %>%
  ggplot(aes(x = sax)) +
  geom_bar() +
  facet_wrap(facets = ~variable)

```
